%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[acmtog]{acmart}
\usepackage[ngerman]{babel}
\usepackage[utf8]{inputenc}

\usepackage{csquotes}
\usepackage{listings}

\lstset{
    language=Python,
    basicstyle=\ttfamily\footnotesize,
    keywordstyle=\color{blue}\bfseries,
    stringstyle=\color{gray},
    commentstyle=\color{darkgray}\itshape,
    showstringspaces=false,
    numbers=left,
    numberstyle=\tiny\color{gray},
    breaklines=true,
    frame=single,
    captionpos=b,
}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows,external}

%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

\copyrightyear{2025}
\acmYear{2025}
\citestyle{acmauthoryear}

%\usepackage[textsize=tiny]{todonotes}

\usepackage[figurename=Fig.]{caption}
\setcopyright{none}
\makeatletter
\renewcommand{\fnum@figure}{Abb. \thefigure}
\makeatother
\addto\captionsngerman{\renewcommand{\figurename}{Abb.}}
\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10010405.10010406.10010417</concept_id>
       <concept_desc>Applied computing~Enterprise architectures</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10010583.10010750.10010751.10010757</concept_id>
       <concept_desc>Hardware~System-level fault tolerance</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10010520.10010575</concept_id>
       <concept_desc>Computer systems organization~Dependable and fault-tolerant systems and networks</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Applied computing~Enterprise architectures}
\ccsdesc[300]{Hardware~System-level fault tolerance}
\ccsdesc[500]{Computer systems organization~Dependable and fault-tolerant systems and networks}

\keywords{Distributed Computing, High Availability, Resilience}
%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Resilienz und Fehlertoleranz in verteilten Systemen}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Maksym Derhachov}
\authornote{Alle Studierenden trugen zu gleichen Teilen zu dieser Arbeit bei.}
\author{Florian Schmidt}
\authornotemark[1]
\author{Lukas Westholt}
\authornotemark[1]
\affiliation{%
  \institution{Hochschule für Technik, Wirtschaft und Kultur Leipzig (HTWK Leipzig)}
  \streetaddress{Karl-Liebknecht-Str. 132}
  \city{Leipzig}
  %\state{Ohio}
  \country{Deutschland}
  \postcode{04277}
}
%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Derhachov, Schmidt und Westholt}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
%  (Abstract-Länge ist typischerweise 15-25 Zeilen lang, in der PDF-Darstellung)

Diese Arbeit untersucht Strategien und Muster zur Steigerung der Resilienz moderner verteilter Anwendungen,
darunter Circuit Breaker, Retry-Muster und Fallback-Strategien.
Ziel ist es, Systeme gegen Ausfälle und Spitzenbelastungen abzusichern, betriebliche Kosten zu reduzieren
und Nutzerzufriedenheit zu gewährleisten.
Die Analyse umfasst allgemeine Ansätze wie Redundanz und Skalierung sowie konkrete Implementierungen in Python
und Beispiele aus der Industrie.
Abschließend werden zentrale Erkenntnisse zusammengefasst und Handlungsempfehlungen abgeleitet.
\end{abstract}

\maketitle

\input{sections/einleitung}
\input{sections/grundlagen}
\input{sections/hauptteil}
\input{sections/discussion}
\input{sections/zusammenfassung}

%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{../belegarbeit-literatur}

% \input{sections/anhang}

\end{document}
\endinput
%%
%% End of file `sample-acmtog.tex'.
